The Concurrent Calculi Formalisation Benchmark

### 2023-10-17 paper/paper.tex
Fix a number of small things.

Now, following Sangiorgi \& Walker~\cite{picalcbook}, we give rules for the compatible
equivalence relation induced by $\sconga\cdot\cdot$, which we still
write as $\scong\cdot\cdot$: namely, add to reflexivity, symmetry and
transitivity the following condition:
 \begin{mathpar}
   \inferrule[Cong]{\sconga P Q}{\scong {C[P]}{C[Q]}}
 \end{mathpar}

@@

\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
\begin{gather*}
  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \\\label{eq:bisim1} \\
  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'} \label{eq:bisim2}
\end{gather*}
We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.

@@

As with the POPLMark and POPLMark Reloaded challegnes, this challenge does not aim to cover every possible
problem in the area. Our benchmark can be extended in two
dimensions: further expanding and combining the current challenges,
and addressing features not covered by the current challenges.

@@

\vspace{-1mm}%
\paragraph{Addressing new calculi and new features.}
Extending in this dimension
means proposing new challenges covering different features of
process calculi, or different proof mechanisation features.
Some interesting aspects of message passing calculi to explore in
further challenges could be multiparty session types and
choreographies, as their meta-theory includes aspects
that we do not cover here (e.g.: well-formed global types,
projection as a partial function) --- although, dually, some aspects
that we cover here do appear in their theory. Also, one could easily
design a challenge that goes beyond our coinduction challenge, by
exploring different notions of bisimilarity (barbed, weak, etc.),
and trace equivalence.

On the side of proof assistants, the current challenge proposes mechanising the
proofs of theorems. An interesting venue to explore would be to take advantage
of other proof assistant features. For example, proof assistants can often
produce certified code (via code extraction, or compiling and running their
definitions). Additionally, while our current challenge is agnostic about proof
automation, a more specific challenge could be designed with the objective of
automating aspects of the proofs. Finally, a challenge could propose the
integration with other formal reasoning tools, namely model checkers. These
tools are in common use in the field --- and combining automated proofs with
proof assistants offers the potential to ease the path towards larger proofs.

>>>>

Now we give rules for the compatible
equivalence relation induced by $\sconga\cdot\cdot$, which we still
write as $\scong\cdot\cdot$: namely, add to reflexivity, symmetry and
transitivity the following condition:
 \begin{mathpar}
   \inferrule[Cong]{\sconga P Q}{\scong {C[P]}{C[Q]}}
 \end{mathpar}

@@

\emph{Strong barbed bisimilarity}, written \( \sbbisim{}{} \), is the largest symmetric relation such that, whenever \( \sbbisim{P}{Q} \):
\begin{gather}
  \observable{P}{\mu}~\textrm{implies}~\observable{Q}{\mu} \label{eq:bisim1}\\
  \transition{P}{\Atau}{P'}~\textrm{implies}~\transition{Q}{\Atau}{\sbbisim{}{} P'} \label{eq:bisim2}
\end{gather}
We say that a relation is a \emph{strong barbed bisimulation} if it satisfies the conditions given above, but is not necessarily the largest such relation, and that \( P \) and \( Q \) are \emph{strong barbed bisimilar} if \( \sbbisim{P}{Q} \).
Note that, since our systems have potentially infinite behaviors, and strong barbed bisimilarity is defined as the largest relation that satisfies the conditions, bisimulation cannot be defined inductively.

@@

As with the POPLMark and POPLMark Reloaded challenges, this challenge does not aim to cover every possible
problem in the area. Our benchmark can be extended in two
dimensions: further expanding and combining the current challenges,
and addressing features not covered by the current challenges.

@@

\vspace{-1mm}%
\paragraph{Addressing new calculi and new features.}
Extending in this dimension
means proposing new challenges covering different features of
process calculi, or different proof mechanisation features.
Some interesting aspects of message passing calculi to explore in
further challenges could be multiparty session types and
choreographies, as their meta-theory includes aspects
that we do not cover here (e.g.: well-formed global types,
projection as a partial function) --- although some aspects
that we cover here do appear in their theory. Also, one could easily
design a challenge that goes beyond our coinduction challenge, by
exploring different notions of bisimilarity (barbed, weak, etc.),
and trace equivalence.

On the mechanisation side, the present challenge proposes mechanising the
proofs of theorems. An interesting avenue to explore would be to take advantage
of other proof assistant features. For example, proof assistants can often
produce certified code (via code extraction, or compiling and running their
definitions). Additionally, while our current challenge is agnostic about proof
automation, a more specific challenge could be designed with the objective of
automating aspects of the proofs. Finally, a challenge could propose the
integration with other formal reasoning tools, namely model checkers. These
tools are in common use in the field --- and combining automated proofs with
proof assistants offers the potential to ease the path towards larger proofs.
